### 2018-08-31 paper_own_docker/main.tex

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[]{algorithm2e}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{xx.xxx/xxx_x}

% ISBN
\acmISBN{978-1-4503-5933-7/19/04}

%Conference
\acmConference[SAC'19]{ACM SAC Conference}{April 8-12, 2019}{Limassol, Cyprus} 
\acmYear{2019}
\copyrightyear{2019}

\acmArticle{4}
\acmPrice{15.00}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{Attack Graph Generation for Micro-service
	Architecture}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

\author{Amjad Ibrahim}
\affiliation{\institution{Technical University of Munich}}
\email{amjad.ibrahim@tum.de}

\author{Stevica Bozhinoski}
\affiliation{\institution{Technical University of Munich}}
\email{stevica.bozhinoski@tum.de}

\author{Prof. Dr. Alexander Pretschner}
\affiliation{\institution{Technical University of Munich}}
\email{alexander.pretschner@tum.de}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}

\input{./chapters/abstract}

\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{Attack Graph Generaton, Computer Security, Microservices}

\maketitle
\input{./chapters/introduction}
\input{./chapters/background}
\input{./chapters/method}
\input{./chapters/evaluation}
\input{./chapters/related_work}
\input{./chapters/conclusion}
\input{./chapters/future_work}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}