### 2018-09-03 paper_own_docker/chapters/background.tex
Background text updated and new references added.

\section{BACKGROUND}

In this section we start by introducing the concept of microservices, their benefits and security implications (Subsection \ref{chap:microservices}). Afterwards we look into vulnerability scanners as tools to generate vulnerabilities of a single host (Subsection \ref{chap:vulnerability_scanners}). At the end we introduce and formally describe attack graphs as methods to disagnose security weakness of a given system composed of multiple hosts (Subsection \ref{chap:attack_graphs}).

\input{./chapters/background_microservices}
\input{./chapters/vulnerability_scanners}
\input{./chapters/background_attack_graphs}