### 2018-09-04 paper_own_docker/chapters/background.tex
Attack graphs subsection and related work redone, some references added and text mostly corrected with help of correctify

\section{BACKGROUND}

In this section, we start by introducing the concept of microservices, their benefits and security implications (Subsection \ref{chap:microservices}). Afterward, we look into vulnerability scanners as tools that generate vulnerabilities for a single host (Subsection \ref{chap:vulnerability_scanners}). In the end, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts (Subsection \ref{chap:attack_graphs}).

\input{./chapters/background_microservices}
\input{./chapters/vulnerability_scanners}
\input{./chapters/background_attack_graphs}