### 2018-09-04 paper_own_docker/chapters/abstract.tex
Added docker vulnerability statistics links, removed horizonal rules from tables to conform with ACM style and some minor text corrections.

Microservices, in contrast to monolithic systems, provide an architecture that is modular and easily scalable. This advantage has resulted in a rapid increase in usage of microservices in recent years. Despite their rapid increase in popularity, there is a lack of work that focuses of their security aspect. Therefore, in the following paper, we present a novel Breath-first Search (BFS) based method for attack graph generation and security analysis of microservice architectures using Docker and Clair.