### 2018-09-18 paper_own_docker/chapters/background.tex
Todos in the background and minor changes in method and evaluation done

\section{BACKGROUND}
\label{chap:background}

We first start by introducing the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. Afterward in Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools that generate vulnerabilities for a single host. At the end in Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts.

\input{./chapters/background_microservices}
\input{./chapters/vulnerability_scanners}
\input{./chapters/background_attack_graphs}