### 2018-09-18 paper_own_docker/chapters/conclusion.tex
intro draft

\section{CONCLUSIONS AND FUTURE WORK}
\label{chap:conclusion}