### 2018-09-18 paper_own_docker/chapters/abstract.tex
checks here and there

Microservices are increasingly dominating the field of service systems, among their many characteristics, practices are technology heterogeneity and reuse and network interfaces. Therefore, with the increase of utilizing third-party components, the potential vulnerabilities existing in a microservice-based system increase. Based on components dependency, these vulnerabilities lead to exposing critical assets of such systems. Similar problems have been tackled in computer networks communities. In this paper, we propose the utilization of attack graphs as a part of the automated development infrastructure used in microservices-based systems. To that end, we relate microservices to network nodes and automatically generate attack graphs that help practitioners to identify, analyze, and prevent plausible attack paths on their microservice-based container networks. We present a complete solution that can be easily embedded into the continuous delivery systems, and show with real-world use cases its efficiency and scalability. 
%Breath-first Search (BFS) based method for attack graph generation and security analysis of microservice architectures using Docker