### 2018-09-20 paper_own_docker/chapters/background.tex

\section{BACKGROUND}
\label{chap:background}

We start by introducing the concept of microservices, their benefits and security implications in Subsection \ref{chap:microservices}. In Subsection \ref{chap:vulnerability_scanners}, we look into vulnerability scanners as tools to scan a single host for vulnerabilities. In Subsection \ref{chap:attack_graphs}, we introduce and formally describe attack graphs as methods to diagnose security weaknesses of a given system composed of multiple hosts.

\input{./chapters/background_microservices}
\input{./chapters/vulnerability_scanners}
\input{./chapters/background_attack_graphs}