### 2018-10-21 paper_own_docker/chapters/abstract.tex
Abstract and introduction updated based on proof-reading feedback

Microservices, which are typically technologically heterogenous and can be deployed automatically, are increasingly dominating service systems. However, with increased utilization of third-party components distributed as images, the potential vulnerabilities in microservice-based systems increase. Based on component dependency, such vulnerabilities can lead to exposing a system's critical assets. Similar problems have been addressed by the computer networks community. In this paper, we propose utilizing attack graphs in the continuous delivery infrastructure of microservices-based systems.  To that end, we relate microservices to network nodes and automatically generate attack graphs that help practitioners identify, analyze, and prevent plausible attack paths in their microservice-based container networks. We present a complete solution that can be easily embedded in continuous delivery systems and demonstrate its efficiency and scalability based on real-world use cases. 
%Breath-first Search (BFS) based method for attack graph generation and security analysis of microservice architectures using Docker